---
title: "Penguins Analysis"
author: "Jay Ashok Chauhan"
format: 
  html:
    theme: lumen
execute: 
  warning: false
  error: false
---

## Install and load required packages

Firstly, we will install the `palmerpenguins` and `tidyverse` for our analysis using data manipulation, exploration and visualisation for penguins dataset. 
**Copy and paste the below code in your R Console only once!**

```{r}
#| label: install-packages-commands
#| eval: false

install.packages("palmerpenguins")
install.packages("tidyverse")
```

Then load the necessary packages-

```{r}
#| label: setup-project

library(tidyverse)
library(palmerpenguins)
library(knitr)
```

## Introduction to the `palmerpenguins` dataset

This dataset contains data of three different penguin species namely Adelie, Chinstrap and Gentoo penguins at the Palmer Station in the Antarctica, it measures key attributes of the penguins related to their mass, bill, flippers, sex, and their origination which can aid in understanding of how data relationships can be generated.

Let's see how some data looks like:

```{r}
#| label: initial-look-at-data

kable(head(penguins))
```

## Exploring population of different species at the islands

Let's generate a quick plot to see the population of penguin species at the different islands at the Palmer station! 

Before that we will create a subset of the data using some `tidyverse` functions!

```{r}
#| label: population-count-penguins-by-island

# Count penguins for each species and island
penguin_population_by_species_island <- penguins |>
  count(species, island)
```

### Generate bar plot

Let's see the data in a bar plot!

```{r}
#| label: fig-bar-plot-population-by-species-island
#| fig-cap: "Population count of different penguin species on the islands at the Palmer Station"

ggplot(penguin_population_by_species_island, aes(x = island, y = n, fill = species)) +
  geom_col(alpha = 0.8) +
  coord_flip() +
  facet_wrap(~species, ncol = 1) +
  scale_fill_viridis_d()  +
  labs(
    x = "Island",
    y = "Population Count",
    fill = "Penguin Species"
  )
```